Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de~Rham complex over convex domains and domains with shellable triangulations

### 2024-10-08 main.tex
Update on Overleaf.

\usepackage{comment}          % 
\usepackage[colorlinks=true]{hyperref}         % For hypertext links
% Make the hyperlinks look similar to arXiv's
\hypersetup{
    colorlinks=true,
    linkcolor=green,    % Color of internal links (sections)
    citecolor=red,    % Color of citations
    urlcolor=blue,     % Color for URLs
    urlbordercolor={1 1 1},
    pdfborder={0.5 0.5 0}  % Removes the border around the links
}
\usepackage{tikz-3dplot}      % 3d plots 
\usepackage{enumitem}
\setitemize[1]{left=0.0cm}

@@

% \tableofcontents

@@

\subsubsection{Review: General results}

@@

\subsubsection{Review: Analytical constants over convex domains}

@@

\subsubsection{Review: Domains star-shaped with respect to a ball}

@@

\subsubsection{Review: Triangulated domains}

@@

\subsubsection{Objectives: convex domains}

@@

\subsubsection{Objectives: Potentials subject to partial boundary conditions}

@@

\subsubsection{Objective: Main results}

@@

Our main objective remains to find potentials for the differential operators of vector calculus and exterior calculus, such as the curl and divergence, over triangulated domains. 
The operator norms of these potentials will serve as computable Poincar\'e--Friedrichs constants. 
In light of the different approaches iscussed above, we seek constants that can be explicitly computed in terms of the mesh geometry, that do not require solving global finite element problems, and that are independent of the boundary regularity of the domain.
% 2. the gradient construction as blueprint 
% outline the essential features

The sequential construction of potentials for gradient vector fields, as discussed earlier, is well-established in the literature and serves as our conceptual blueprint. 
Gradient potentials are easily computed over each individual simplex, but the constants of integration generally do not match: 
the scalar piecewise potential will belong to a broken Sobolev space. 
In particular, the local potentials will differ only by a constant along the simplex boundaries. 
However, we can produce a potential in Sobolev spaces over increasingly larger intermediate subdomains, which we construct sequentially. 
At each step, we select a simplex that shares a face with one of the previously processed simplices and adjust the constant of integration of the local potential. 
The global potential is built cell by cell until the entire domain is covered. 
Our main result in this context is Theorem~\ref{theorem:poincarefriedrichsestimate}.

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation} if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that triangulation.\footnote{Simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes~(cf.\ \cite{ziegler1995lectures}) \cye{and simply ``simplicial meshes'' in the finite element literature}.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, 
which is the union $\underlying{\calT} = \bigcup \calT$ of all simplices in $\calT$.
% \begin{align*}
    % \underlying{\calT} := \bigcup_{ T \in \calT } T
    % .
% \end{align*}
We call any set triangulable if it is the underlying set of some triangulation.

@@

\begin{remark}
    In three dimensions, 
    the calculus of differential forms is in correspondence with classical vector calculus. 
    This is expressed formally as the commuting diagram 
    \begin{gather*}
    \begin{CD}
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) @>{\cartan}>> C^\infty\Alt^3(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        @VV{\varpi^{3}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{3} @>{\curl}>> C^\infty(\Omega)^{3} @>{\divergence}>> C^\infty(\Omega)
    \end{CD}
    \cye{\, ,}
    \end{gather*}
    where $\varpi^{0}$ and $\varpi^{3}$ are the identity mappings and where 
    \begin{align*}
     \varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} + u_3 \cartanx^{3} \right) 
     &= 
     \left( u_1, u_2, u_3 \right), 
     \\
     \varpi^{2}\left( u_{12} \cartanx^{1} \wedge \cartanx^{2} + u_{13} \cartanx^{1} \wedge \cartanx^{3} + u_{23} \cartanx^{2} \wedge \cartanx^{3} \right) 
     &= 
     \left( u_{23}, - u_{13}, u_{12} \right).   
    \end{align*}
    In two dimensions, 
    the calculus of differential forms can be translated into 2D vector calculus in two different ways. 
    To the authors' best knowledge, neither convention is dominant over the other in the literature.
    We summarize the situation in the following commuting diagram: 
    \begin{gather*} 
    \begin{CD}
        C^\infty(\Omega) @>{\curl}>> C^\infty(\Omega)^{2} @>{\divergence}>> C^\infty(\Omega)
        \\
        @AA{\varkappa^{0}}A 
        @AA{\varkappa^{1}}A 
        @AA{\varkappa^{2}}A 
        \\
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{2} @>{\rot}>> C^\infty(\Omega)
    \end{CD}
    \cye{\,.}
    \end{gather*}
    Here, $\varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} \right) = \left( u_1, u_2 \right)$ is the lower middle isomorphism. We introduce the rotation operator $J(x,y) = (y,-x)$ and define $\varkappa = J \varpi$ and $\rot = \divergence J$.
    The other vertical arrows are the identity. 
    The utility of exterior calculus is that the operators of vector calculus can be translated into a common framework that does not depend on the dimension.
\end{remark}

>>>>

\usepackage{xcolor}          % 
\usepackage{comment}          % 
\usepackage[colorlinks=true]{hyperref}         % For hypertext links
% Make the hyperlinks look similar to arXiv's
\hypersetup{
    colorlinks=true,
    linkcolor=Maroon,    % Color of internal links (sections)
    citecolor=red,    % Color of citations
    urlcolor=blue,     % Color for URLs
    urlbordercolor={1 1 1},
    pdfborder={0.5 0.5 0}  % Removes the border around the links
}
\usepackage{tikz-3dplot}      % 3d plots 
\usepackage{enumitem}
\setitemize[1]{left=0.0cm}

@@

\tableofcontents

@@

\subsubsection{General results}

@@

\subsubsection{Analytical constants over convex domains}

@@

\subsubsection{Domains star-shaped with respect to a ball}

@@

\subsubsection{Triangulated domains}

@@

\subsubsection{Convex domains}

@@

\subsubsection{Potentials subject to partial boundary conditions}

@@

\subsubsection{Main results}

@@

Our main objective remains to find potentials for the differential operators of vector calculus and exterior calculus, such as the curl and divergence, over triangulated domains. 
The operator norms of these potentials will serve as computable Poincar\'e--Friedrichs constants. 
In light of the different approaches iscussed above, we seek constants that can be explicitly computed in terms of the mesh geometry, that do not require solving global finite element problems, and that are independent of the boundary regularity of the domain.

% 2. the gradient construction as blueprint 
% outline the essential features

The sequential construction of potentials for gradient vector fields, as discussed earlier, is well-established in the literature and serves as our conceptual blueprint. 
Gradient potentials are easily computed over each individual simplex, but the constants of integration generally do not match: 
the scalar piecewise potential will belong to a broken Sobolev space. 
In particular, the local potentials will differ only by a constant along the simplex boundaries. 
However, we can produce a potential in Sobolev spaces over increasingly larger intermediate subdomains, which we construct sequentially. 
At each step, we select a simplex that shares a face with one of the previously processed simplices and adjust the constant of integration of the local potential. 
The global potential is built cell by cell until the entire domain is covered. 
Our main result in this context is Theorem~\ref{theorem:poincarefriedrichsestimate:grad}.

@@

A finite family of simplices $\calT$ is a \emph{simplicial complex} or \emph{triangulation} if it satisfies the following conditions: 
(i) $\calT$ contains all the subsimplices of its members (ii) any non-empty intersection of two members of $\calT$ is a common subsimplex of each of them. 
We say that a simplicial complex $\calT$ has \textit{dimension $n$} or is \textit{$n$-dimensional} if each of its simplices is a subset of an $n$-dimensional member of that triangulation.\footnote{Simplicial complexes that we call $n$-dimensional are called purely $n$-dimensional in the literature on polytopes~(cf.\ \cite{ziegler1995lectures}) and simply ``simplicial meshes'' in the finite element literature.} 
We also write $\underlying{\calT}$ for the \textit{underlying set} of the simplicial complex $\calT$, 
which is the union $\underlying{\calT} = \bigcup \calT$ of all simplices in $\calT$.
% \begin{align*}
    % \underlying{\calT} := \bigcup_{ T \in \calT } T
    % .
% \end{align*}
We call any set triangulable if it is the underlying set of some triangulation.

@@

\begin{remark}
    In three dimensions, 
    the calculus of differential forms is in correspondence with classical vector calculus. 
    This is expressed formally as the commuting diagram 
    \begin{gather*}
    \begin{CD}
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) @>{\cartan}>> C^\infty\Alt^3(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        @VV{\varpi^{3}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{3} @>{\curl}>> C^\infty(\Omega)^{3} @>{\divergence}>> C^\infty(\Omega)
    \end{CD}
    \, ,
    \end{gather*}
    where $\varpi^{0}$ and $\varpi^{3}$ are the identity mappings and where 
    \begin{align*}
     \varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} + u_3 \cartanx^{3} \right) 
     &= 
     \left( u_1, u_2, u_3 \right), 
     \\
     \varpi^{2}\left( u_{12} \cartanx^{1} \wedge \cartanx^{2} + u_{13} \cartanx^{1} \wedge \cartanx^{3} + u_{23} \cartanx^{2} \wedge \cartanx^{3} \right) 
     &= 
     \left( u_{23}, - u_{13}, u_{12} \right).   
    \end{align*}
    In two dimensions, 
    the calculus of differential forms can be translated into 2D vector calculus in two different ways. 
    To the authors' best knowledge, neither convention is dominant over the other in the literature.
    We summarize the situation in the following commuting diagram: 
    \begin{gather*} 
    \begin{CD}
        C^\infty(\Omega) @>{\curl}>> C^\infty(\Omega)^{2} @>{\divergence}>> C^\infty(\Omega)
        \\
        @AA{\varkappa^{0}}A 
        @AA{\varkappa^{1}}A 
        @AA{\varkappa^{2}}A 
        \\
        C^\infty\Alt^0(\Omega) @>{\cartan}>> C^\infty\Alt^1(\Omega) @>{\cartan}>> C^\infty\Alt^2(\Omega) 
        \\
        @VV{\varpi^{0}}V 
        @VV{\varpi^{1}}V 
        @VV{\varpi^{2}}V 
        \\
        C^\infty(\Omega) @>{\grad}>> C^\infty(\Omega)^{2} @>{\rot}>> C^\infty(\Omega)
    \end{CD}
    \,.
    \end{gather*}
    Here, $\varpi^{1}\left( u_1 \cartanx^{1} + u_2 \cartanx^{2} \right) = \left( u_1, u_2 \right)$ is the lower middle isomorphism. We introduce the rotation operator $J(x,y) = (y,-x)$ and define $\varkappa = J \varpi$ and $\rot = \divergence J$.
    The other vertical arrows are the identity. 
    The utility of exterior calculus is that the operators of vector calculus can be translated into a common framework that does not depend on the dimension.
\end{remark}
